Spatial phylogenetics of Japanese ferns: Patterns, processes, and implications for conservation

### 2022-01-18 ms/manuscript.Rmd
Update MS
Follow AJB format

Short running title: Spatial phylogenetics of Japanese ferns

Joel H. Nitta^1^, Brent D. Mishler^2^, Wataru Iwasaki^1,3^, and Atsushi Ebihara^4^

@@

Author for correspondence:
*Joel H. Nitta*
*Email: joelnitta@gmail.com*

## Acknowledgements

Members of the Iwasaki Lab at the University of Tokyo provided comments that improved the manuscript.
Shawn Laffan provided advice about CANAPE.
This study was supported by JSPS KAKENHI Grant Number 16H06279.

@@

### Study Site

@@

### Occurrence data

@@

We further cleaned the list prior to analysis by filtering out any occurrences not within the second-degree mesh and removing duplicate collections (`r occ_sum$filtered$n_specimens$value %>% scales::number(big.mark = ",")` specimens, `r occ_sum$filtered$n_taxa$value` taxa after filtering).
Given the high quality of our occurrence data and the fact that automated occurrence cleaning algorithms [`r eg`, @Zizka2019] have the potential to erroneously exclude true occurrence points (`r ie`, false positives), we chose not to apply additional automated cleaning steps to our data as is often done with occurrence records obtained from GBIF [`r eg`, @Rice2019a; @Suissa2021].

@@

Besides habitat loss due to human activity, one major threat to ferns in Japan is herbivory by Japanese deer (*Cervus nippon*).
Although native, Japanee deer have caused extensive damage to native plant communities due to rapid population growth and range expansion since the 1970s [@Takatsuki2009].
Decrease or extirpation of fern populations due to deer herbivory has been observed in multiple sites and species across the country [@Minamitani2005; @Yahara2006; @Hattori2010].
Furthermore, 33 of 212 (15.5%) fern taxa listed as endangered (Red List status CR, EN or VU) in Japan include herbivory by deer as one of the reasons for their endangered status, and one taxon (*Arachniodes yasu-inouei* Sa. Kurata var. *angustipinnula* Seriz.) is thought to have gone extinct due to deer herbivory, among other causes [@MOJ2015; @kankyosho2022].
Although not all species of ferns are equally susceptible to deer herbivory [a small number are even apparently unpalatable to deer\; @Minamitani2005], these studies show it is a clear danger to many species.
Therefore, analysis of the threat posed to fern biodiversity by deer herbivory is needed to inform future conservation policies.

@@

## Results

### Datasets

@@

## Discussion

### Distribution and drivers of biodiversity

@@

### Limits of the study and future outlook

@@

# Acknowledgements

@@

# Author contributions

@@

# Data availability statement

@@

# Online supporting information

@@

# Literature cited
<!-- Insert <div id="refs"></div> to place the refs here instead of at the end -->
<div id="refs"></div>

@@

**`r figure("conserv-status")`**. Conservation status and threat due to herbivory by deer in areas with high levels of biodiversity for ferns of Japan. (a) Percent of land area with medium or high protected status for grid-cells with significantly high biodiversity by protection status. (b) Percent of land area occupied by deer for grid-cells with significantly high biodiversity by deer range data source (1978 survey, 2003 survey, or range estimated from model based on 2003 survey data). Biodiversity metrics include taxon richness, phylogenetic diversity (PD), functional diversity (FD), and phylogenetic endemism (PE). Significance of PD, FD, and PE assessed by a one-tailed test comparing observed values to a null distribution of 999 random values; for richness, grid-cells in the top 5% considered highly diverse (see Materials and Methods). For (a), vertical lines indicate total percent area protected across Japan (baseline protection rate). For (b), vertical lines indicate total percentage of land area occupied by deer across Japan (baseline threat rate).

`r pagebreak_doc()`

```{r japan-map-doc, fig.cap = "", fig.width = 5, fig.height = 5, eval = params$doc_type == "doc", results = "asis"}
# Print plot if rendering to doc on initial submission
japan_map_plot
cat("\n\n", figure("japan-map"))
```

`r pagebreak_doc()`

```{r raw-diversity-doc, fig.cap = "", fig.width = 7, fig.height = 4, eval = params$doc_type == "doc", results = "asis"}
raw_diversity_plot
cat("\n\n", figure("raw-diversity"))
```

`r pagebreak_doc()`

```{r rand-diversity-doc, fig.cap = "", fig.width = 7, fig.height = 9, eval = params$doc_type == "doc", results = "asis"}
rand_diversity_plot
cat("\n\n", figure("rand-diversity"))
```

`r pagebreak_doc()`

```{r coeff-doc, fig.cap = "", fig.width = 7, fig.height = 5, eval = params$doc_type == "doc", results = "asis"}
coeff_plot
cat("\n\n", figure("coeff"))
```

`r pagebreak_doc()`

```{r modelfit-doc, fig.cap = "", fig.width = 7, fig.height = 5, eval = params$doc_type == "doc", results = "asis"}
modelfit_plot
cat("\n\n", figure("modelfit"))
```

`r pagebreak_doc()`

```{r endemism-doc, fig.cap = "", fig.width = 5, fig.height = 5, eval = params$doc_type == "doc", results = "asis"}
endemism_plot
cat("\n\n", figure("endemism"))
```

`r pagebreak_doc()`

```{r bioregions-doc, fig.cap = "", fig.width = 7, fig.height = 5, eval = params$doc_type == "doc", results = "asis"}
bioregions_plot
cat("\n\n", figure("bioregions"))
```

`r pagebreak_doc()`

```{r div-by-tax-region-doc, fig.cap = "", fig.width = 7, fig.height = 7, eval = params$doc_type == "doc", results = "asis"}
div_by_tax_region_plot
cat("\n\n", figure("div-by-tax-region"))
```

`r pagebreak_doc()`

```{r conserv-status-doc, fig.cap = "", fig.width = 7, fig.height = 4, eval = params$doc_type == "doc", results = "asis"}
conserv_status_plot
cat("\n\n", figure("conserv-status"))
```

`r pagebreak_doc()`

## Data availability statement

Code to replicate all analyses, figures, and this manuscript are available at https://github.com/joelnitta/japan_ferns_spatial_phy.
A Docker image to run the code is available at https://hub.docker.com/r/joelnitta/japan_ferns_spatial_phy.
Data files needed to run the analysis and selected results files are available from Dryad at http://doi.org/FIXME:ADD_DOI_HERE_WHEN_ASSIGNED.

## Biosketch

Joel H. Nitta's research focuses on patterns and processes of diversification in ferns. For more information, see https://joelnitta.com.

Author contributions: J.H.N., B.M., and A.E. conceived the ideas; A.E. provided the data; J.H.N. analyzed the data; W.I provided resources; J.H.N. wrote the manuscript with input from all co-authors.

>>>>

Joel H. Nitta^1,5^, Brent D. Mishler^2^, Wataru Iwasaki^1,3^, and Atsushi Ebihara^4^

@@

^5^Author for correspondence: joelnitta@gmail.com

```{r author-page-doc, results = "asis", echo = FALSE, eval = params$doc_type == "docx"}
# Author page, version for submission to journal
cat('
Manuscript received _______; revision accepted _______.

Short running title: Spatial phylogenetics of Japanese ferns
')
```

@@

## Study Site

@@

## Occurrence data

@@

We further cleaned the list prior to analysis by filtering out any occurrences not within the second-degree mesh and removing duplicate collections (`r occ_sum$filtered$n_specimens$value %>% scales::number(big.mark = ",")` specimens, `r occ_sum$filtered$n_taxa$value` taxa after filtering).
Given the high quality of our occurrence data and the fact that automated occurrence cleaning algorithms [`r eg`, CoordinateCleaner\; @Zizka2019] have the potential to erroneously exclude true occurrence points (`r ie`, false positives), we chose not to apply additional automated cleaning steps to our data as is often done with occurrence records obtained from GBIF [`r eg`, @Rice2019a; @Suissa2021].

@@

Besides habitat loss due to human activity, one major threat to ferns in Japan is herbivory by Japanese deer (*Cervus nippon*).
Although native, Japanese deer have caused extensive damage to native plant communities due to rapid population growth and range expansion since the 1970s [@Takatsuki2009].
Decrease or extirpation of fern populations due to deer herbivory has been observed in multiple sites and species across the country [@Minamitani2005; @Yahara2006; @Hattori2010].
Furthermore, 33 of 212 (15.5%) fern taxa listed as endangered (Red List status CR, EN or VU) in Japan include herbivory by deer as one of the reasons for their endangered status, and one taxon (*Arachniodes yasu-inouei* Sa. Kurata var. *angustipinnula* Seriz.) is thought to have gone extinct due to deer herbivory, among other causes [@MOJ2015; @kankyosho2022].
Although not all species of ferns are equally susceptible to deer herbivory [a small number are apparently unpalatable to deer\; @Minamitani2005], these studies show it is a clear danger to many species.
Therefore, analysis of the threat posed to fern biodiversity by deer herbivory is needed to inform future conservation policies.

@@

# RESULTS

## Datasets

@@

# DISCUSSION

## Distribution and drivers of biodiversity

@@

## Limits of the study and future outlook

@@

## ACKNOWLEDGEMENTS

@@

## AUTHOR CONTRIBUTIONS

@@

## DATA AVAILABILITY STATEMENT

@@

## SUPPORTING INFORMATION

@@

@@

**`r figure("conserv-status")`**. Conservation status and threat due to herbivory by deer in areas with high levels of biodiversity for ferns of Japan. (A) Percent of land area with medium or high protected status for grid-cells with significantly high biodiversity by protection status. (B) Percent of land area occupied by deer for grid-cells with significantly high biodiversity by deer range data source (1978 survey, 2003 survey, or range estimated from model based on 2003 survey data). Biodiversity metrics include taxon richness, phylogenetic diversity (PD), functional diversity (FD), and phylogenetic endemism (PE). Significance of PD, FD, and PE assessed by a one-tailed test comparing observed values to a null distribution of 999 random values; for richness, grid-cells in the top 5% considered highly diverse (see Materials and Methods). For (A), vertical lines indicate total percent area protected across Japan (baseline protection rate). For (B), vertical lines indicate total percentage of land area occupied by deer across Japan (baseline threat rate).
